### 2020-12-11 main.tex
AWR body submission

@@

@@

\subsection{Layers}

\begin{enumerate}
\item Problem -> symbol embeddings
\item Symbol embedding -> symbol cost
\item Symbol costs -> precedence cost
\end{enumerate}

@@

We define precedence cost:
$$
\CostPrec(\pi) =
C \sum_{1 \leq i \leq n} \CostSym(\pi(i)) \cdot i =
C \sum_{1 \leq i \leq n} \CostSym(s_i) \cdot \inv{\pi}(s_i)
$$
Precedence cost is minimized by $\pi$ that orders the symbols by their costs in non-decreasing order.

@@

We define precedence pair cost:
$$
\CostPrecPair(\pi_0, \pi_1) =
\CostPrec(\pi_0) - \CostPrec(\pi_1) =
C \sum_{1 \leq i \leq n} \CostSym(s_i) \cdot [\inv{\pi_0}(s_i) - \inv{\pi_1}(s_i)]
$$
Clearly $\CostPrecPair(\pi_0, \pi_1) < 0$ iff $\CostPrec(\pi_0) < \CostPrec(\pi_1)$.
For a pair of precedences about which we know that $\pi_0$ is better than $\pi_1$,
we want $\CostPrecPair(\pi_0, \pi_1) < 0$.

We model the probability of misclassification of precedence pair $\pi_0, \pi_1$
such that $\pi_0$ is better than $\pi_1$
by the sigmoid of $\CostPrecPair(\pi_0, \pi_1)$:
$$
L(\pi_0, \pi_1) = \sigmoid(\CostPrecPair(\pi_0, \pi_1))
$$

>>>>

\usepackage[pdf]{graphviz}

@@

\vampire{} \cite{10.1007/978-3-642-39799-8_1}
is representative
of the state-of-the-art \gls{fol} \gls{atp} design.
\vampire{} was awarded the first place
in the \gls{fol} category of \gls{casc} \cite{},
an annual international competition of \glspl{atp},
in eleven times in the years between 2010 and 2020 \cite{}.
Moreover, 7 of 10 provers that have placed in the top 50 \% in 2020
are saturation-based, refutation-based and resolution-based,
similarly to \vampire{} \cite{}.

%\vampire{} is a saturation-based prover.
%Given a \gls{fol} problem,
%\vampire{} converts the problem into \gls{cnf}.
%and proceeds to saturate the set of clauses
%by repeated application of inference rules.

%\vampire{} searches for a proof by applying the rules
%of the superposition inference system
%\cite{10.1007/978-3-642-39799-8_1}.

% AWR candidate
Proof search in \vampire{} is crucially constrained
by a simplification ordering on terms.
Being refutation-based and saturation-based,
\vampire{} searches for contradiction
by iteratively inferring clauses provable from the input clauses.
The inferences

explores the space of clauses provable
from the input formulas
by applying rules of superposition inference system.
This process continues until a contradiction is inferred,
the processed clauses set is saturated
or the execution hits a resource limit.

During proof search,
\vampire{} explores the space of provable clauses
by applying inference rules to clauses
that have already been proven.

% AWR candidate
Simplification ordering on terms
influences the proof search in \vampire{} on two levels.
First, the inferences on each clause are limited
to the selected literals.
In each clause,
either a negative literal or all the maximal literals are selected.
The maximality is evaluated
according to the simplification ordering.
% Note: The selection function Total does not use the simplification ordering.
Second, simplification ordering orients some of the equalities
to prevent superposition and equality factoring
from inferring redundant complex conclusions.
In each of these two roles,
the simplification ordering may impact the direction and,
in effect, the length of the proof search.

The inference system used in the state-of-the-art \gls{fol} \gls{atp} \vampire{}
is parameterized by a simplification ordering on terms.
\vampire{} uses superposition inference system \cite{}.

The \gls{atp} \vampire{} uses superposition inference system.

Vampire supports two simplification term ordering classes:
\gls{lpo} and \gls{kbo}.

The superposition inference system \cite{} used in Vampire is parameterized
by symbol precedence.

% Example: It seems that a classical example for KBO an orienting equations is group theory axioms.

The choice of symbol precedence may affect the length of a proof search to a great extent.

Given a \gls{fol} problem,
determining a symbol precedence that leads to a fast proof search is a non-trivial task.

@@

This section describes the architecture of a precedence recommender system.
When presented with a \gls{fol} problem $P$ with symbols $\symbols$,
the system predicts a symbol precedence $\pi$ on $\symbols$
that is expected to lead to a successful proof search.

\subsection{Prediction pipeline}

\subsection{Training pipeline}

\begin{figure}[ht]
\caption{Architecture overview}
\centering
\digraph[scale=0.4]{precedencepairclassifierdetailed}{
	graph [splines=ortho];
	node [shape=renctangle, fontsize=20];
	edge [fontsize=20];
	fol [label="FOL problem", shape=oval];
	pi0 [shape=oval, label=<&pi;<SUB>0</SUB>>];
	pi1 [shape=oval, label=<&pi;<SUB>1</SUB>>];
	invpi0 [label=<&pi;<SUB>0</SUB><SUP>-1</SUP>>];
	invpi1 [label=<&pi;<SUB>1</SUB><SUP>-1</SUP>>];
	cnf [label="Clause normal form (CNF)"];
	symbolembeddings [label="Symbol embeddings"];
	symbolcosts [label="Symbol costs"];
	pi1pi0 [label="Inverse precedence difference"];
	normalized [label="Normalized inverse precedence difference"];
	paircost [label="Precedence pair cost"];
	fol -> cnf [xlabel=" Vampire "];
	cnf -> symbolembeddings [xlabel=< <B>Graph Convolution Network</B> >];
	symbolembeddings -> symbolcosts [xlabel=< <B>Feed-forward neural network</B> >, style=bold];
	symbolcosts -> paircost [style=bold];
	paircost -> loss [xlabel=" Binary cross-entropy ", style=bold];
	loss [label="Loss", shape=oval];
	pi0 -> invpi0 [xlabel=" Invert "];
	pi1 -> invpi1 [xlabel=" Invert "];
	invpi0 -> pi1pi0;
	invpi1 -> pi1pi0;
	pi1pi0 -> normalized [xlabel=" Normalize "];
	normalized -> paircost;
	symbolprecedence [label="Symbol precedence", style=dashed];
	symbolcosts -> symbolprecedence [xlabel=" Order symbols by their costs ", style=dashed];
}
\end{figure}

\subsection{Layers}
'
\begin{enumerate}
\item Problem -> symbol embeddings
\item Symbol embedding -> symbol cost
\item Symbol costs -> precedence cost
\end{enumerate}

@@

We define precedence cost:
$$
\CostPrec(\pi) =
C \sum_{1 \leq i \leq n} \CostSym(\pi(i)) \cdot i =
C \sum_{1 \leq i \leq n} \CostSym(s_i) \cdot \inv{\pi}(s_i)
$$
Precedence cost is minimized by $\pi$ that orders the symbols by their costs in non-increasing order
($\forall (1 \leq i < j \leq n) . (\CostSym(\pi(i)) \geq \CostSym(\pi(j)))$).

@@

% Note that we use this orientation because the TensorFlow metric BinaryCrossentropy classifies 0 as negative and we use the value 0 for "failed to classify" logits.
Given a pair of precedences $\pi_0, \pi_1$,
we define the log-odds of the event "$\pi_0$ is better than $\pi_1$":
$$
\CostPrecPair(\pi_0, \pi_1) =
\CostPrec(\pi_1) - \CostPrec(\pi_0) =
C \sum_{1 \leq i \leq n} \CostSym(s_i) \cdot [\inv{\pi_1}(s_i) - \inv{\pi_0}(s_i)]
$$
Clearly $\CostPrecPair(\pi_0, \pi_1) > 0$ iff $\CostPrec(\pi_0) < \CostPrec(\pi_1)$.
For a pair of precedences about which we know that $\pi_0$ is better than $\pi_1$,
we want $\CostPrecPair(\pi_0, \pi_1) > 0$.

We model the probability of the event "$\pi_0$ is better than $\pi_1$"
by the sigmoid of $\CostPrecPair(\pi_0, \pi_1)$:https://www.overleaf.com/project/5f75a50ba1a0930001ce1162
$$
p(\pi_0, \pi_1) = \sigmoid(\CostPrecPair(\pi_0, \pi_1))
$$

We use the binary cross-entropy loss to train the model.
Given a pair of precedences such that $\pi_0$ is better than $\pi_1$,
the loss is as follows:
$$
Loss(\pi_0, \pi_1) = -\log(\sigmoid(\CostPrecPair(\pi_0, \pi_1)))
$$
